%%

\section{Assignment: Ten myths about FLOSS business models}


\begin{frame}
\frametitle{Dynamics and goals}

\begin{itemize}
\item Statements based on ``Ten myths about FLOSS business models'', by Carlo Daffara \\
\url{http://www.groklaw.net/articlebasic.php?story=20070828132340846}
\item Each one can be right or wrong (or arguable).
\item Work in groups to give simple answers supporting or rebating the statement.
\item Document your answers with factual arguments, if possible (links, data, etc.).
\end{itemize}
\end{frame}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\begin{frame}
 \frametitle{Statement 1}
 \begin{center}
  \begin{LARGE} FLOSS does not prevent prices from being established for it.  \end{LARGE}
 \end{center}

\end{frame}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\begin{frame}
 \frametitle{Statement 2}
 \begin{center}
  \begin{LARGE} FLOSS licenses aim to suppress any ownership claims, they are hostile to
intellectual property rights.  \end{LARGE}
 \end{center}

\end{frame}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\begin{frame}
 \frametitle{Statement 3}
 \begin{center}
  \begin{LARGE} FLOSS development is mainly driven by ad-hoc altruism and volunteer effort. \end{LARGE}
 \end{center}

\end{frame}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\begin{frame}
 \frametitle{Statement 4}
 \begin{center}
  \begin{LARGE} If I release my software to the FLOSS community, thousands of developers 
will suddenly start working for me, for nothing. \end{LARGE}
 \end{center}

\end{frame}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\begin{frame}
 \frametitle{Statement 5}
 \begin{center}
  \begin{LARGE} Many big companies and firms (even outside software development business) use FLOSS. \end{LARGE}
 \end{center}

\end{frame}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\begin{frame}
 \frametitle{Statement 6}
 \begin{center}
  \begin{LARGE} FLOSS is inherently unreliable, and not very well supported. \end{LARGE}
 \end{center}

\end{frame}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\begin{frame}
 \frametitle{Statement 7}
 \begin{center}
  \begin{LARGE} Private companies contribute to FLOSS, since they can get profit in exchange. \end{LARGE}
 \end{center}

\end{frame}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\begin{frame}
 \frametitle{Statement 8}
 \begin{center}
  \begin{LARGE} Widespread FLOSS projects are invariably funded by large private companies. \end{LARGE}
 \end{center}

\end{frame}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\begin{frame}
 \frametitle{Statement 9}
 \begin{center}
  \begin{LARGE} FLOSS software has achieved to become a market leader in certain sectors.  \end{LARGE}
 \end{center}

\end{frame}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\begin{frame}
 \frametitle{Statement 10}
 \begin{center}
  \begin{LARGE} If a company leading a FLOSS project is acquired, the new owner
can lock the source code and force all its customers to pay for it. \end{LARGE}
 \end{center}

\end{frame}

