%%

\section{Activity: the open core debate}

\begin{frame}
\frametitle{Introduction}
\begin{itemize}
  \item \textit{Open core} is one of the frequent, yet most controversial
business model strategies around FLOSS.
  \item Open core companies defend their right to maintain proprietary licenses
on strategic features and modules.
  \item FLOSS advocates argue that this approach takes benefit from FLOSS without
giving substantial feedback to communities and projects.
\end{itemize}
\end{frame}

\begin{frame}
 \frametitle{References}
  \begin{enumerate}
   \item Lampitt. ``Open core licensing...''
   \texttt{\footnotesize{http://alampitt.typepad.com/lampitt\_or\_leave\_it/}}
   \texttt{\footnotesize{2008/08/open-core-licen.html}}
   \item Carlo Daffara. ``Relationships between open core, dual licensing and contributions''
   \texttt{\footnotesize{http://carlodaffara.conecta.it/?p=460}}
   \item Simon Phipps. ``Open core is bad for you''.
   \texttt{\footnotesize{http://blogs.computerworlduk.com/simon-says/2010/06/}}
   \texttt{\footnotesize{open-core-is-bad-for-you/}}
   \item OSI. ``A simple declaration about open core''
   \texttt{\footnotesize{http://www.opensource.org/blog/OpenCore}}
  \end{enumerate}

\end{frame}

\begin{frame}
 \frametitle{Questions}
 \begin{itemize}
  \item Find 3 examples of companies following open core strategies.
  \item Could any of these companies switch to a different business model
  and ensure sustainability? How?
  \item Do these companies make any claims on this issue? Summarize
  their arguments.
  \item After learning the arguments from both sides of the story, what is your
  own opinion on this issue?
 \end{itemize}

\end{frame}



