%%

\section{Single-vendor vs. Community owned FLOSS}

\begin{frame}
\frametitle{Introduction}
\begin{itemize}
  \item 
  \end{itemize}
\end{itemize}
\end{frame}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\begin{frame}
\frametitle{Questions}
\begin{itemize}
  \item Create groups of 3-4 persons to address the following challenges.
  \item Remember to substantiate your answers with adequate references.
  \begin{itemize}
   \item Find out one example of company identified with each of the 4 stages presented.
   \item What are the main control points and steering mechanisms associated to each stage?
   \item Compare the case of Mozilla Foundation projects and RedHat's Fedora. Explain the
common and different control points and steering mechanisms adopted by each community.
  \end{itemize}
\end{itemize}
\end{frame}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\begin{frame}
\frametitle{References}
\begin{itemize}
  \item The Single-Vendor Commercial Open Source Business Model, by D. Riehle.
  \begin{itemize}
   \item \footnotetext{\textit{http://dirkriehle.com/publications/2009/}}
   \footnotetext{\textit{the-commercial-open-source-business-model/}}
  \end{itemize}
  \item The Economic Case for Open Source Foundations, by D. Riehle.
  \begin{itemize}
   \item \footnotezie{\textit{http://dirkriehle.com/publications/2010/}}
   \footnotezie{\textit{the-economic-case-for-open-source-foundations/}}
  \end{itemize}
  \item The Economic Motivation of Open Source Software: Stakeholder Perspectives, by D. Riehle
  \begin{itemize}
   \item \footnotezie{\textit{http://dirkriehle.com/computer-science/research/2007/}}
   \footnotezie{\textit{computer-2007-article.html}}
  \end{itemize}
  \item The future of open source licensing, by The 451 Group.(Matthew Aslett).
  \item http://blogs.the451group.com/opensource/2010/08/25/the-future-of-open-source-licensing/
   \item Estimating savings from OSS code reuse, or: where does the money comes from?, by C. Daffara.
  \begin{itemize}
   \item \footnotetext{\textit{http://carlodaffara.conecta.it/?p=135}}
  \end{itemize}

  \end{itemize}
\end{itemize}
\end{frame}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
