\documentclass[a4paper]{article}
%\usepackage[T1]{fontenc}
\usepackage[utf8]{inputenc}
\usepackage{url}
\usepackage[hypertex,colorlinks]{hyperref}
%\usepackage{html}
%\usepackage{hthtml}
\usepackage{geometry}

% Comments (optional argument is author of comment)
\newcommand{\comments}[2][?]{
  \begin{quote}
    \textbf{Comment (#1):} {\em #2}
  \end{quote}
  }

% \name is for ``special names'', like procedure or variable names.
\newcommand{\name}[1]{\texttt{\hturl{#1}}}

% Just a shortcut for links where the url appears as footnote.
\newcommand{\htflink}[2]{\htmladdnormallinkfoot{#1}{#2}}

\title{MSWL Master Thesis \\
Master on libre software \\
URJC - GSyC/Libresoft \\
\url{http://master.libresoft.es}}

\author{Gregorio Robles}
\date{January 2011}

\sloppy
\begin{document}
\maketitle

\begin{abstract}
Course syllabus and learning program for the course ``Master Thesis'', of the Master on libre software of the Universidad Rey Juan Carlos (Móstoles, Spain).

[This is an evolving document, until the course is finished and graded]
\end{abstract}

\tableofcontents

\section{Grading}

This section details the criteria for grading the course, the deadlines for the different activities, and the submission details for th activities that require them.

\subsection{Evaluation criteria}
\label{sub:evaluation-criteria}

This subject will be evaluated considering following criteria:

\begin{itemize}

\item A complete and personal master thesis dissertation that documents the efforts undertaken by the student during this subject. The master thesis dissertation must follow
the official rules of the Universidad Rey Juan Carlos and the ones, if they exist, of the Master on Libre Software, which can both be found from the URJC and master web site.

\item The defense of the master thesis. To accomplish this issue, a committee will be set up that will analyze the master thesis dissertation and ask students to creat a 3 to 10 minute video that presents the most important facts and conclusions of the master thesis. The committee can ask the student for enhancements and corrections. The committee can be composed of professors, lecturers of the master and any other developer involved in the libre software project that is related the master thesis done by the student.

\end{itemize}

\subsection{Submission deadlines}



\end{document}
