%% presentation.tex
%%
%% Presentation of the course ``Legal Aspects'' of the Official Master on Libre Software (URJC)
%% http://master.libresoft.es
%%

%%---------------------------------------------------------------------
%%---------------------------------------------------------------------

\section{Presentation of the course}

%%---------------------------------------------------------------

\begin{frame}
\frametitle{Administrative data}

\begin{itemize}
\item First semester, 3 ECTS credits
\item Teachers:
  \begin{itemize}
  \item \alert{Miguel Vidal} (\texttt{mvidal at gsyc.urjc.es})
  \item \alert{Gregorio Robles} (\texttt{grex at gsyc.urjc.es})
  \end{itemize}
\item Schedule: \alert{Thursdays, 17:00-21:00}
  \begin{itemize}
  \item Classroom 227, Departamental I, campus at Fuenlabrada
%  \item Lab 7, Laboratorios II, campus at Fuenlabrada
  \end{itemize}
\item Moodle course: \\
  {\footnotesize\url{http://gsyc.escet.urjc.es/moodle/course/view.php?id=121}}
\end{itemize}
\end{frame}


%%---------------------------------------------------------------

\begin{frame}
\frametitle{Subject: Legal Aspects}

\setbeamercovered{invisible}

\center \Large \textbf{Why should I care about legal aspects?}
 
\pause

\bigskip

\begin{center}
\alert{Licenses are legal basis and ``social contract'' of free/open source software.}
\end{center}

\end{frame}

%%---------------------------------------------------------------

\begin{frame}
\frametitle{Main topics addressed}

\begin{itemize}
   \item Intellectual property (IP) law \alert{for nonlawyers}: Copyrights,
Patents, Trademarks, Warranties...
   \item How FLOSS and IP work together
   \item Free and Open Source Software licenses (permissive and reciprocal)
   \item Free cultural works (documentation, images, Creative Commons
licenses)
   \item Case studies and practical issues (choosing a license, compatibility,
forking, merging/mixing... myths)
\end{itemize}

\end{frame}


%%---------------------------------------------------------------

\begin{frame}
\frametitle{Sessions}

\begin{enumerate}
\item \alert{Oct. 13:} Introduction to the general concepts of Intellectual Property
\item \alert{Oct. 20:} Copyright on Software. Types of FLOSS licenses
\item \alert{Oct. 27:} Free Software Licenses
\item \alert{Nov. 3:} Free licenses for other intellectual works
\item \alert{Nov. 10:} Case studies and myths
\end{enumerate}

\end{frame}


%%---------------------------------------------------------------

\begin{frame}
\frametitle{Evaluation criteria}

\begin{itemize}
\item \alert{Specific report}: \textit{45 points}
\item \alert{Exercises} (in Moodle forum): \textit{35 points}
\item \alert{Blog entries} (optional): \textit{20 points} 
\item \alert{Other activities} (optional, agreed with lecturers): \textit{15 points} 
\end{itemize}

\end{frame}

%%---------------------------------------------------------------

\begin{frame}
\frametitle{Grades}

\begin{itemize}
\item \textbf{0-49: Fail} (``Suspenso'', SS)
\item \textbf{50-69: Pass} (``Aprobado'', AP)
\item \textbf{70-89: Good} (``Notable'', NT)
\item \textbf{90-100: Excellent} (``Sobresaliente'', SB)
\end{itemize}

\end{frame}


%%---------------------------------------------------------------

\begin{frame}
\frametitle{Final work: Specific Report}

\begin{itemize}
\item Specific written report about a relevant legal aspect of free/open source software. 
\item Mandatory.
\item About 7-10 pages long.
\item Deadline: December 23rd.
\end{itemize}

\end{frame}

%%---------------------------------------------------------------

\begin{frame}
\frametitle{Final work: Some ideas}

\begin{itemize}
\item IP (copyrights, patents) lawsuits against FLOSS.
\item Infringement and violations of FLOSS licenses.
\item How to affect SaaS (software on demand, cloud computing, etc.) to FLOSS licensing.
\item Analysis of a specific FLOSS license or comparison between FLOSS licenses.
\item Thinking about derivate works: linking and licensing, types of business models with FLOSS licenses, etc.
\item Intellectual Property debate: Defense vs. opposition against patents and copyrights.
\end{itemize}

\end{frame}


%%---------------------------------------------------------------


\begin{frame}
\frametitle{Some references}

\begin{itemize}
\item Course Website: \\
  {\footnotesize \url{http://gsyc.escet.urjc.es/moodle/course/view.php?id=121}}
\item \textsc{Van Lindberg}, \textit{Intellectual Property and Open Source}, O'Reilly, July 2008.
\item \textsc{Malcolm Bain} et al. \textit{Aspectos legales y de explotación del software libre}, UOC, February 2007. \\
{\footnotesize \url{http://ocw.uoc.edu/informatica-tecnologia-y-multimedia/aspectos-legales-y-de-explotacion-del-software-libre/materiales/}}
\item \textsc{Lawrence Rose}, \textit{Open Source Licensing}, Prentice Hall, July 2004 
\end{itemize}

\end{frame}



